---
title: "TEST"
author: "Erika Vargas"
date: "May 19, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#load data sorted by year and month to reflect the first sale of each producer/agent
library(readxl)
DATAORG <- read_excel("~/Desktop/SPRING 2019/CAPSTONE/DATAORG.xlsx")
head(DATAORG)
```

```{r}
# select distinct producers (eliminating duplicate)
unique_producers <- subset(DATAORG, !duplicated(`Producer SalesForce Contact ID`))
nrow(unique_producers)  #number of distinct agents/producers
```

```{r}
# loading dataset that contains the y varibale "0", "1"
# where 1 represents the first sale and 0 that it wasnt the first sale
firstsale <- read_excel("salesDistinct.xlsx")
class(firstsale)
names(firstsale)[1] = c("Producer SalesForce Contact ID")
str(firstsale)
```

```{r}
#joining tables
data <- merge(x = unique_producers, y = firstsale, by = "Producer SalesForce Contact ID", all = TRUE)
head(data)
```

#cleaning the data
```{r}
data <- data[,-c(4,13,14,15,16,17,18,19,20,21)] #removes an empty and redundant columns
str(data)

#renaming columns
names(data)[1:14] =
  c("ProducerID","Ext_Channel", "Ext_Specialist", "Parent", "FirmName", "State", "PSuiteCT", "ProductName", "Year" , "Month", "FirstSaleAmount","Y2016","Y2017","Y2018")

#since most of the variables are categorical I must convert them into factors
library(aod)

#changing character columns to factors 
data <- mutate_at(data, vars(Ext_Channel,Ext_Specialist,Parent,FirmName,State,PSuiteCT, ProductName,Year, Month, Y2016,Y2017, Y2018), as.factor)

str(data)
summary(data)

#delete rows that contain NA
data <- na.omit(data)
summary(data)
str(data)
nrow(data) # number of agents we are going to use for the analysis 
```

*Before looking at the first sale amount per agent, I want to see the top 10 producers/agents and the total amount of sales they produced*
```{r}
#total sales amount by producer
SalesByProducer <- aggregate(`Production Amount`~`Producer SalesForce Contact ID`,data=DATAORG,FUN=sum)
# top 10 Producers
SalesByProducer %>%top_n(10)
# Bottom 10 Producers
SalesByProducer %>%top_n(-10)
```

```{r}

```


#2016 Analysis
```{r}
#amount of the first sale by year 2016-2018
sales2016_ <- subset(data, Year == 2016)

#looking at amount of the first sale by Month in 2016
boxplot(log(FirstSaleAmount)~ Month,data=sales2016_, main="Sales in 2016", 
   xlab="Month", ylab=" First Sale Amount")

# Load function
source("http://pcwww.liv.ac.uk/~william/R/crosstab.r")

#row percentages for sales by month
crosstab(data, row.vars = "Y2016", col.vars = "Month", type = "r")

#row percentages for external channel first sellers
crosstab(data, row.vars = "Y2016", col.vars = "Ext_Channel", type = "r")

#row percentages for PSuite CT first sellers 
crosstab(data, row.vars = "Y2016", col.vars = "PSuiteCT", type = "r")

#row percentages for state first sellers 
crosstab(data, row.vars = "Y2016", col.vars = "State", type = "r")
```

*Most of the first sales in 2016 happened in April, February, March, May. The monthS with less first sales are August, December, July, June, November, October September. Therefore, I can see that most of the first sales in 2016 happened in the first semester of the year.*

*63.16% of first sales in 2016 used a bank as an external channel, 26.42% used a wirehouse, and 10.42% used an independent external channel*

*61.46% of first sales in 2016 were FIA products, followed by 31.02%  for fixed Annuities, 6.61% for IA& ANNUITIZATIONS, 0% for RIA,and 0.90% for variable annuities*

*California seems to be the state where most of the first sales (13.82%) in 2016 took place, followed by florida (8.38%), New York (8.38%),Texas (5.85%), OH (5.23%), PA (5.26%)*


```{r}
#LOGIT REGRESSION
#Logistic regression, also called a logit model, is used to model dichotomous outcome variables, I n this case weather the produce had a first sale or not (1 or 0). In the logit model the log odds of the outcome is modeled as a linear combination of the predictor variables.

mylogit <- glm(Y2016 ~ Ext_Channel + Ext_Specialist + Parent + FirmName + State + PSuiteCT + ProductName + Month + FirstSaleAmount, data = data, family = "binomial")

summary(mylogit)
```

