---
title: "salesData"
author: "Erika Vargas"
date: "May 4, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#loading the data
library(dplyr)
library(magrittr)
library(readxl)
sales_data <- read_excel("symetra_data.xlsx")
head(sales_data)
```

```{r}
#cleaning the data
sales_data <- sales_data[,-c(4,13,14,15)] #removes an empty column
class(sales_data) # the data is a dataframe
#renaming columns
names(sales_data)[1:11] =
  c("Ext_Channel", "Ext_Specialist", "ProducerID", "Parent", "Firm Name", "State", "PSuite CT", "ProductName", "Year" , "Month", "SalesAmount")

head(sales_data)
str(sales_data)
```

```{r}
#total unique producers 
total_producers <- subset(sales_data, !duplicated(ProducerID))
nrow(total_producers)
```

#looking at Sales by producers, Top 10.
```{r}
#total sales amount by producer
SalesByProducer <- aggregate(SalesAmount~ProducerID,data=sales_data,FUN=sum)
# top 10 Producers
SalesByProducer %>%top_n(10)
# Bottom 10 Producers
SalesByProducer %>%top_n(-10)
```

```{r}
#amount of sales by year 2016-2018

sales2016 <- subset(total_producers, Year == 2016)
sales2017 <- subset(total_producers, Year == 2017)
sales2018 <- subset(total_producers, Year == 2018)

#looking at sales by Month in 2016
boxplot(log(SalesAmount)~ Month,data=sales2016, main="Sales in 2016", 
   xlab="Month", ylab="Sales Amount")

#looking at sales by Month in 2017
boxplot(log(SalesAmount)~ Month,data=sales2017, main="Sales in 2017", 
   xlab="Month", ylab="Sales Amount")

#looking at sales by Month in 2018
boxplot(log(SalesAmount)~ Month,data=sales2018, main="Sales in 2018", 
   xlab="Month", ylab="Sales Amount")


```


```{r}
#import data with sales by year 
ProducerSalesByyear <- read_excel("salesDistinct.xlsx")
class(ProducerSalesByyear)
names(ProducerSalesByyear)[1] = c("ProducerID")
str(ProducerSalesByyear)

```


```{r}
#joining tables to consolidate the data
data <- merge(x = total_producers, y = ProducerSalesByyear, by = "ProducerID", all = TRUE)
head(data)
# I will delete the 2019 column because of the values are N/A
data <- data[,-17]
```


```{r}
library(aod)
library(ggplot2)

#changing character columns to factors 
data_clean <- mutate_at(data, vars(Ext_Channel,Ext_Specialist,Parent,`Firm Name`,State,`PSuite CT`, ProductName,Year, Month, `2016 First Sale`, `2017 First Sale`, `2018 First Sale`), as.factor)

str(data_clean)
summary(data_clean)
```

```{r}
#first sellers in 2016 (1)
table(data_clean$`2016 First Sale`)

#first sellers in 2017 (1)
table(data_clean$`2017 First Sale`)

#first sellers in 2018 (1)
table(data_clean$`2018 First Sale`)

```


```{r}
#checking for empty sales using crosstab
# Load function
source("http://pcwww.liv.ac.uk/~william/R/crosstab.r")
# Frequency count
crosstab(data_clean, row.vars = "2016 First Sale", col.vars = "Ext_Channel", type = "f")

#row percentages for external channel first sellers
crosstab(data_clean, row.vars = "2016 First Sale", col.vars = "Ext_Channel", type = "r")

#63.43% of the first sales in 2016 use a bank as an external channel, 26.18% used a wirehouse, and 10.369% used a independent external channel 


#row percentages for PSuite CT first sellers 
crosstab(data_clean, row.vars = "2016 First Sale", col.vars = "PSuite CT", type = "r")

#68.46% of first sales in 2016 were FIA products, followed by 25.69% of Annuities 

crosstab(data_clean, row.vars = "2016 First Sale", col.vars = "Month", type = "r")
# most of the first sales in 2016 happened in April, February, March, May. The month with less first sales are August, December, July, June, November, October September. therefore we can see that most of the first sales happen in the first semester of teh year. 
```

#LOGIT REGRESSION
Logistic regression, also called a logit model, is used to model dichotomous outcome variables, I n this case weather the produce had a first sale or not (1 or 0). In the logit model the log odds of the outcome is modeled as a linear combination of the predictor variables.

```{r}
mylogit <- glm(`2016 First Sale` ~ Ext_Channel + Ext_Specialist + `PSuite CT` + Month + `2016`, data = data_clean, family = "binomial")

summary(mylogit)
```

This fitted model says that, holding all predictor variable at a fixed value, the odds of having a first sale into using independent as external channel ()over the oods of of having a first sale using wirhouse as external channel) is exp(0.815) = 0.664.  


```{r}
#We can test for an overall effect of external channel using the wald.test function of the aod library
#for external channel
wald.test(b = coef(mylogit), Sigma = vcov(mylogit), Terms = 2:3)

#for ext especialist
wald.test(b = coef(mylogit), Sigma = vcov(mylogit), Terms = 4:37)

# for psuite
#for external channel
wald.test(b = coef(mylogit), Sigma = vcov(mylogit), Terms = 38:40)

# for month
wald.test(b = coef(mylogit), Sigma = vcov(mylogit), Terms = 41:51)
```

```{r}
## odds ratios and 95% CI
exp(cbind(OR = coef(mylogit), confint(mylogit)))
```


#### Linear Probaility Model 
P(y=1|x) = B0 + B1X1 + .... +BkXk
```{r}
#changing character columns to factors 
data_clean <- mutate_at(data, vars(Ext_Channel,Ext_Specialist,Parent,`Firm Name`,State,`PSuite CT`, ProductName,Year, Month, `2016 First Sale`, `2017 First Sale`, `2018 First Sale`), as.factor)

str(data_clean)

model1 <- lm(`2016 First Sale` ~ Ext_Channel + Ext_Specialist + Parent + State + `PSuite CT` + ProductName + Month + `2016`, data = data_clean)
summary(model1)
```



















#GROUPING BY YEARS
**I would like to look at the data by years and see if I found correlations between factors for that first sale**


```{r}
# producer with first sale in 2016
one2016 <- data %>%
        filter(!is.na(data$`2016`), data$`2016`!=0, data$`2016` != "-" , data$Year == 2016,
               is.na(data$`2015`), is.na(data$`2014`), is.na(data$`2017`), is.na(data$`2018`))

str(one2016)
#changing character columns to factors 
one2016 <- mutate_at(one2016, vars(Ext_Channel,Ext_Specialist,Parent,`Firm Name`,State,`PSuite CT`, ProductName,Year, Month), as.factor)

```

```{r}
#2016
#contigency tables
parent_Firmname <- table(one2016$`Firm Name`, one2016$Parent)
table(one2016$Ext_Specialist,one2016$Ext_Channel)
table(one2016$`PSuite CT`,one2016$Ext_Channel)

#checking for independency using chi-square test 
tbl = table(one2016$`PSuite CT`,one2016$Ext_Channel)
chi2 <- chisq.test(tbl)
chi2$statisti
chi2$p.value

# computing V. the smaller v, the lower the correlation)
sqrt(chi2$statistic / sum(tbl))

#ANALYZING numercial vs categorical variable using anova table 

fit <- aov(one2016$`2016` ~ one2016$State, data = one2016)
plot(fit)
```



```{r}
#visuals of the data
library(ggfortify)

### look what kind of principal component analysis can be done with categorical data

boxplot(log(`2016`)~ Ext_Channel,data=one2016, main="Sales by External Channel in 2016", 
   xlab="Ext.Channel", ylab="Sales Amount")

hist(log(one2016$`2016`), breaks = 50)

```




```{r}
#looking for possible predictors 
#state
regression2014 <- lm(log(`2014`) ~ State, data= one2014)
summary(regression2014)
plot(regression2014)
```













```{r}
#removing blank rows in production amount 

 sales_data[!(is.na(sales_data$SalesAmount) | sales_data$SalesAmount==""), ]
```

